---
title: "R"
author: 李朋飞
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: zh-CN.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
setwd("/Users/lpf/LPF/bitbucket/data-science/WIKI/program/R")
```
# 简介

## The R Project for Statistical Computing


R is a free software environment for **statistical computing** and **graphics**

R is ‘GNU S’, a freely available language and environment for statistical computing and graphics which provides a wide variety of statistical and graphical techniques: linear and nonlinear modelling, statistical tests, time series analysis, classification, clustering, etc

.footnote[
1. https://cloud.r-project.org/index.html
]

???
R是一个统计和绘图软件

---
# 特点

1. 包含了大量的统计方面的计算

2. 处理数据量小

3. 画图功能强大

4. 扩展性强(包)

# 目录

1. R语言安装与简介

2. R语言基本数据结构及操作

3. R语言数据读写 

4. R包

  4.1 data.table
  
  4.2 RMySQL
  
  4.3 reshape2
  
  4.4 ggplot2
  
  4.5 shiny


---
# 1. R语言安装与入门

## R 安装

- Windows (exe)

- MAC (DMG)

- LINUX 

```
# install R
yum install R
```

.footnote[
1. https://cloud.r-project.org/index.html
2. https://www.rstudio.com/products/rstudio/download/
]

---
## IDE : Rstudio 

RStudio is a set of integrated tools designed to help you be more productive with R. It includes a console, syntax-highlighting editor that supports direct code execution, and a variety of robust tools for plotting, viewing history, debugging and managing your workspace.

- install Rstudio
```
$ wget https://download2.rstudio.org/rstudio-server-rhel-1.1.447-x86_64.rpm
$ sudo yum install rstudio-server-rhel-1.1.447-x86_64.rpm
```
- 访问 url : ip:8787

---
## R 入门

```{r}
age = c(1,3,5,2,11,9)
weight = c(4.4,5.3,7.2,5,8.5,7.3)
mean(weight)
sd(weight)
cor(age,weight)
```
---
```{r}
par(family="STHeiti")
plot(age,weight,main = "体重随年龄的变化")
```



---
# 2. R语言数据结构

- 向量
- 矩阵
- 数据框
- 列表
- 因子
- 数组

---
## 数据类型
- 字符
```{r}
"a"
```
- 数字
```{r}
1
```
- 逻辑变量
```{r}
c(T,F,TRUE,FALSE)
```
---
- 空
```{r}
NA    ## 占位符
length(NA)
NULL  ## 空
length(NULL)
```

---
## 向量
```{r}
a = 1 
print(a)
b = c(1,2,3,4,-1,0)
print(b)
d = c("a","b","c")
print(d)
e = c(TRUE,FALSE,T,F)
print(e)
```
---
## 向量特点

1. 同一数据类型

2. 非同一数据类型会进行转换

```{r}
( a = c("a",1,3) )
( b = c(1,TRUE) )
( d = c(T,"a") )
```
---
## 向量下标

```{r}
a = c("a","b","c","d")
length(a)
a[c(1,3)]
a[1:3]
a[-1]
```

---
## 矩阵

- 数据元素类型唯一

```{r}
a = matrix(1:10,nrow=2,dimnames = c())
a
a[2]
a[,1]
a[1,]
```
---
## 矩阵 --> 向量

```{r}
vector_a = as.vector(a)
vector_a
class(vector_a)
is.vector(a)
```

---
## 向量 --> 矩阵

```{r}
matrix_a = as.matrix(vector_a)
matrix_a
matrix(vector_a,nrow=2)
```
---
## 数据框

- 每一列的元素类型可以不同

```{r}
df = data.frame(
  name = letters[1:6],
  age = c(1,3,5,2,11,9),
  weight = c(4.4,5.3,7.2,5,8.5,7.3)
)
df
```
---
### 数据框属性
```{r}
dim(df)
colnames(df)
rownames(df)
str(df)
```

---
### 数据框操作
- 下标
```{r}
df[c(1,2),c(1:3)]
df[c(1,2),c("name","age")]
df[1,]  # 取某行 
df[,2] # 取某列
```
---
- 返回值类型

```{r}

df1 = df[2]
df1
class(df1)
df2 = df[,2]
df2
class(df2)
```

---
- 返回值类型
```{r}
df3 = df[,2,drop=FALSE]
df3
class(df3)
```

---
### 数据框合并

```{r}
df3 = cbind(df,df)
df3 
df4 = rbind(df,df)
df4
```

---
## 列表

```{r}
a = c(1:3)
b = c ("a","b","c","d")
e = matrix(1:10,nrow=2)
l = list(a=a,b=b,e=e)
l
```
---
## 列表基本操作
```{r}
l[1]
l['a']  # 返回list
l[['a']] # 返回存储内容
l$a      # 返回存储内容
```

---
## 因子
```{r}
factor(letters[c(1:5,1:2)], labels = "letter")
(m = factor(2:3))
levels(m)
levels(m)=1:2
m
```
---
- 因子类型转换
```{r}
as.numeric(m)
as.numeric(as.character(m))  # 因子变为字符串，再变成数字
```

---
## 控制和函数
- for, if, while
```{r}
x = letters
len = length(x)
for (i in 1:len) {
  if(i>20) {
    print (paste(i,x[i]))
  }
}
```

---
## 函数
```{r}
print_max = function(x){
  max_value = max(x)
  print(sprintf("the max value is : %s",max_value))
  return(max_value)
}
print_max
x = c(4,2,8,10)
max_x = print_max(x)
max_x
```

---
# 文件读写
- 数据框 : 规则数据
```{r}
f = tempfile()
d = data.frame(a=sample(10),b=letters[1:10])
write.table(d,file=f,    # 写文件
            row.names = FALSE,
            col.names = TRUE,
            sep="\t"
            )

e = read.table(f,header = T, # 读文件
               sep="\t",
               stringsAsFactors = FALSE
               )
head(e)

```

---
## 其他常用操作
- 日期格式数据

```{r}
d = c("2018-08-01","2018-08-02")
class(d)
d_date = as.Date(d)
class(d_date)
d_date+3
d_date[1] - d_date[2]  # as.numeric转为numeric
```

---
- merge
```{r}
d = data.frame(date=Sys.Date()-1:10,
              users = sample(10),
              group=letters[1:10],
              stringsAsFactors = FALSE
              )
head(d)
str(d)
```
---
- merge
```{r}
d1 = data.frame(group=letters[1:7],money=sample(7)*100)
head(d1)
d2 = merge(d,d1,by="group",all.x=TRUE)
d2
```
---
- order : 排序
```{r}
d2[order(d2$money),]
```
- subset ：子集
```{r}
subset(d2,users>6&money>200)
# d2[d2$users>6&d2$money>200,]
```
---
- apply 分组计算
```{}
apply # 可以按行或者按列进行计算
sapply # 对list中每个元素进行计算
tapply # 分组计算
```

---
- 正则
```{r,eval=FALSE}
gsub  # 全局替换
grep  # 返回匹配的索引
grepl # 返回TRUE和FALSE
```

---
## 赋值
```{r}
a = 1:3
a <- 1:3
```
---
## 查看帮助
```{r,eval=FALSE}
?lapply
?"for"                  # but quotes/backticks are needed
?`?`
?women                  # information about data set "women"
example(plot)
```
---
## 常用函数
```
c : 生成向量
matrix : 生成矩阵
data.frame : 生成数据框
list : 生成列表
as.data.frame : 变为数据框
as.character : 变为字符
as.numeric : 变为数据
class : 查看数据类型
str : 查看数据结构
length : 长度
nrow : 数据框/矩阵 行数
ncol : 数据框/矩阵 列数
```
---
## 常用函数
```
sum : 求和
min : 最小值
mean : 均值
sd : 方差
Sys.Date : 获取日期
as.Date : 字符串转换日期
strftime : 时间格式化
strptime : 字符串变为日期
sort : 排序
order : 排序值
paste : 合并
subset : 取子集
unique : 去重
coSums : 按列求和
rowSums : 按行求和
```
---
# R包

Currently, the CRAN package repository features 13022 available packages.


```{r}
# .packages(all.available=TRUE) #列出可用R包
# install.packages("ggplot2") # 安装包
# library(ggplot2)  # 加载包
# library(help="ggplot2") # 查看包所有函数
# ?`ggplot2-package`   # 包的帮助文档
search() #查看已经加载的R包
```
.footnote[
- ref

[CRAN Task Views](https://cran.r-project.org/web/views/)
]


---
# 连接数据库 RMySQL

```{r,echo=FALSE}
talk = list(
  host = "47.93.15.159",
  port = 52111,
  user = "data",
  password = "Data_0723"
)
```
```{r}
library(RMySQL)
db <- dbConnect(
    MySQL(),  host = talk$host,
    port = talk$port, user = talk$user,
    password = talk$password
  )
data <- dbGetQuery(db,"select * from talk.user limit 10 ")
head(data)
```
---
- 建立连接之后，进行数据查询，之后最好断开连接，因为超过16个连接之后会报错，在shiny中要特别注意。
```{r}
dbDisconnect(db) # 断开连接
# dbWriteTable
# library(help=RMySQL)
```

---
# data.table 

继承与data.frame,但是更强大，更方便

```{r}
library(data.table)
## 读文件
dt = fread(f) # fast than read.table and more convient
dt
class(dt)
```
---
- 数据库转为data.table
```{r}
DF = data.frame(x=rep(c("b","a","c"),each=3), y=c(1,3,6), v=1:9)
class(DF)
setDT(DF)
class(DF)
```
---
- 子集

```{r,eval=FALSE}
DT[2]                          # 2nd row
DT[3:2]                        # 3rd and 2nd row
DT[order(x)]                   # no need for order(DT$x)
DT[order(x), ]                 # same as above. The ',' is optional
DT[y>2]                        # all rows where DT$y > 2
DT[y>2 & v>5]                  # compound logical expressions
DT[!2:4]                       # all rows other than 2:4
DT[-(2:4)]                     # same
```
---
- 分组计算


```{r}
DT = DF
head(DT)
DT[, sum(v), by=x]             # ad hoc by, order of groups preserved in result
DT[, sum(v), keyby=x]          # same, but order the result on by cols
```
---
- 分组计算
```{r}
DT[,.(sum = sum(v) , avg = mean(v)),keyby = list(x,y)]
DT[,.(sum = sum(v) , avg = mean(v)),keyby = list(x)]

```

---
- 添加新列

```{r}
DT[,sum_v := sum(v)]
DT
DT[,sum_v_byx := sum(v),by=x]
DT
```
---
- 其他

```{r}
DT[,.N]    # 行数
DT[,.N,by = x ] # 分组计算行数
DT[,.SD[1:2]]   # 前两行
# DT[,.SD[1:2], by = x]   # 分组前两行
DT[,.SD[,1:2]]
```

---
# reshape2
- melt

```{r}
require(reshape2)
names(airquality) <- tolower(names(airquality))
head(airquality)
melt(airquality, id=c("month", "day"))
```
---
- acast
```{r}
names(airquality) <- tolower(names(airquality))
aqm <- melt(airquality, id=c("month", "day"), na.rm=TRUE)
head(aqm)
acast(aqm, month ~ variable, mean)
```

---
# ggplot2
- 点图
```{r}
require(ggplot2)
p <- ggplot(mtcars, aes(x = wt, y = mpg))
p + geom_point()
```
---
- 线图
```{r}
p+geom_line(color = "red") +labs(x = "x轴",y="y轴",title = "标题 : 这是个线图") + 
  theme(text = element_text(family = "STHeiti"), # 设置中文字体(MAC)
        plot.title = element_text(hjust = 0.5) #标题居中
        ) 
 
```
---
```{r}
p + geom_line() + geom_point() +labs(x = "x轴",y="y轴",title = "标题 : 这是个线+点图") + 
  theme(text = element_text(family = "STHeiti"),# 设置中文字体
        plot.title = element_text(hjust = 0.5) #标题居中
        ) 
```
---
- 分组线图
```{r}
head(mtcars)
```

```{r,eval = FALSE}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(aes(colour = factor(cyl), shape = factor(vs))) +
  labs(
    x = "Weight (1000 lbs)",
    y = "Fuel economy (mpg)",
    colour = "Cylinders",
    shape = "Transmission"
   )
```
---
```{r,echo=FALSE}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(aes(colour = factor(cyl), shape = factor(vs))) +
  labs(
    x = "Weight (1000 lbs)",
    y = "Fuel economy (mpg)",
    colour = "Cylinders",
    shape = "Transmission"
   )
```

---
- 柱状图
```{r}
g <- ggplot(mpg, aes(class))
# Number of cars in each class:
g + geom_bar()
```
---
```{r}
g +
 geom_bar(aes(fill = drv), 
          position = position_stack(reverse = F)) + 
 coord_flip() + 
 theme(legend.position = "top")
```
---
- position 参数说明


 名称 | 描述
 ---- | ----
 dodge  | 避免重叠，并排放置
 fill | 堆叠元素并将高度标准化为1
 identify | 不做任何调整
 jitter | 给点添加扰动避免重合
 stack  | 将图形元素堆叠起来


.footnote[
-ggplot2：数据分析与图形艺术
]

---
- pie 

---
- 可交互画图包
```{r}
# recharts  : echart的R语言接口
# plotly    : 可以与ggplot完美结合，将ggplot2的图形转换为可交互图形
g = ggplot(mtcars,  aes(wt, mpg)) + geom_point()
plotly::ggplotly(g)
```
---
# 相关参考资料

1. R example 
2. R语言实战
3. R语言编程艺术
4. 官网文档

